---
title: "Demo on sentimental analysis"
output: html_document
---

```{r}
library(tidyverse)
library(tidytext)
library(janeaustenr)
```


In the previous demo, we have tidied the work of Jane Austen.

```{r}
original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()
tidy_books <- original_books %>% 
  unnest_tokens(word, text) %>%
  mutate(word = str_extract(word, "[a-z]*")) %>%
  filter(word != "")
```

# Simple sentiment analysis 

Sentiment analysis is one of the most obvious things we can do with unlabelled text data (with no score or no rating) to extract some insights out of it. One of the most primative sentiment analysis is to perform a simple dictionary lookup and calculate a final composite score based on the number of occurences of positive and negative words. In this section, we will explore this approach with Jane Austen's works.

The `tidytext` package contains four sentiment lexicons in the sentiments dataset, three general-purpose ones and one suitable for financial text (e.g., where "share" is not necessarily positive and "liability" not necessarily negative). 

The three general-purpose lexicons are

- `AFINN` from [Finn Arup Nielsen](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010),
- `bing` from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html), and
- `nrc` from [Saif Mohammad and Peter Turney](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm).

All three of these lexicons are based on unigrams, i.e., single words and can be accessed via the function `get_sentiments()`.

```{r}
library(textdata)
get_sentiments("afinn") 
get_sentiments("afinn") %>% count(value)

get_sentiments("bing")
get_sentiments("bing") %>% count(sentiment)

get_sentiments("nrc")
get_sentiments("nrc") %>% count(sentiment)
```

**Question:** Take a look at the first few records in each lexicons and explore how the sentimental scores are organized.

With data in a tidy text format, sentiment analysis can be done as an inner join. Let's look at the words with a joy score from the NRC lexicon and find the most common joy words in *Emma*.

```{r}
nrc_joy <- get_sentiments("nrc") %>% filter(sentiment == "joy")

tidy_books %>% filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}
nrc_anger <- get_sentiments("nrc") %>% filter(sentiment == "anger")

tidy_books %>% filter(book == "Emma") %>%
  inner_join(nrc_anger) %>%
  count(word, sort = TRUE)
```


We see mostly positive, happy words about hope, friendship, and love here. Furthermore, we can also examine how sentiment changes throughout each novel. First, We find a sentiment score for each word using the Bing lexicon. Then, we count up how many positive and negative words there are in defined sections of each book. 

Note that small sections of text may not have enough words in them to get a good estimate of sentiment while really large sections can wash out narrative structure. For these books, using 80 lines works well, but this can vary depending on individual texts, how long the lines were to start with, etc.

```{r}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  mutate(index = linenumber %/% 80 ) %>%
  count(book, index, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

Now we can plot these sentiment scores across the plot trajectory of each novel. Notice that we are plotting against the index on the x-axis that keeps track of narrative time in sections of text.

```{r}
ggplot(jane_austen_sentiment) +
  geom_col(aes(index, sentiment, fill = book), show.legend = FALSE) +
  facet_wrap( ~ book, ncol = 2, scales = "free_x")
```


**Question:** With several options for sentiment lexicons, you might want some more information on which one is appropriate for your purposes. Use all three sentiment lexicons and examine how the sentiment changes across the narrative arc of Pride and Prejudice. 

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

modified_nrc <- get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive","negative"))

pride_prejudice %>%
  inner_join(modified_nrc) %>%
  mutate(index = linenumber %/% 80 ) %>%
  count(index, sentiment) %>%
  spread(key=sentiment, value = n, fill = 0) %>% 
  mutate(score = positive - negative) -> nrc_novel_sentiment

ggplot(nrc_novel_sentiment) +
  geom_col(mapping = aes(x=index,y=score))
```

# Your codes goes here

## afinn
```{r} 
get_sentiments("afinn")%>%
  inner_join(pride_prejudice) %>%
  mutate(index = linenumber %/% 80 ) %>%
  group_by(index) %>% 
  summarise(score = mean(value)) -> afinn_novel_sentiment
```
```{r}
ggplot(afinn_novel_sentiment) + geom_col(mapping = aes(x=index,y=score))
```

